---
title: "R : Comparing Oversampling and Undersampling Algorithms for Class Imbalance Treatment"
author: "John Pauline Pineda"
date: "April 29, 2023"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
    highlight: tango
    css: doc.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=15, fig.height=10)
```

# **1. Table of Contents**

##  1.1 Sample Data

```{r section_1.1, warning=FALSE, message=FALSE}
```

##  1.2 Data Quality Assessment

```{r section_1.2, warning=FALSE, message=FALSE}
```

##  1.3 Data Preprocessing

###  1.3.1 Outlier

```{r section_1.3.1, warning=FALSE, message=FALSE}
```

###  1.3.2 Zero and Near-Zero Variance

```{r section_1.3.2, warning=FALSE, message=FALSE}
```

###  1.3.3 Collinearity

```{r section_1.3.3, warning=FALSE, message=FALSE}
```

###  1.3.4 Linear Dependencies

```{r section_1.3.4, warning=FALSE, message=FALSE}
```

###  1.3.5 Shape Transformation

```{r section_1.3.5, warning=FALSE, message=FALSE}
```

###  1.3.6 Centering and Scaling

```{r section_1.3.6, warning=FALSE, message=FALSE}
```

###  1.3.7 Pre-Processed Dataset

```{r section_1.3.7, warning=FALSE, message=FALSE}
```

## 1.4 Data Exploration

```{r section_1.4, warning=FALSE, message=FALSE}
```

## 1.5 Oversampling and Undersampling Algorithms Applied for Class Imbalance using Logistic Regression

###  1.5.1 Original Data (REF)

```{r section_1.5.1, warning=FALSE, message=FALSE}
```

###  1.5.2 Oversampling - Adaptive Synthetic Algorithm (OS_ADASYN)

```{r section_1.5.2, warning=FALSE, message=FALSE}
```

###  1.5.3 Oversampling - Borderline Synthetic Minority Oversampling Technique  (OS_BSMOTE)

```{r section_1.5.3, warning=FALSE, message=FALSE}
```

###  1.5.4 Oversampling - Synthetic Minority Oversampling Technique  (OS_SMOTE)

```{r section_1.5.4, warning=FALSE, message=FALSE}
```

###  1.5.5 Oversampling - Random Oversampling Examples  (OS_ROSE)

```{r section_1.5.5, warning=FALSE, message=FALSE}
```

###  1.5.6 Undersampling - Near Miss Algorithm  (US_NM)

```{r section_1.5.6, warning=FALSE, message=FALSE}
```

###  1.5.7 Undersampling - Tomek Links  (US_TL)

```{r section_1.5.7, warning=FALSE, message=FALSE}
```

## 1.6 Evaluation Summary

```{r section_1.6, warning=FALSE, message=FALSE}
```

# **2. References**